### 2024-04-09 latex/draft.tex
Pushing initial latex creation.

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}

\title{Causal Inference with Graphical Neural Networks}
\author{Your Name}
\date{\today}
\maketitle

\section{Introduction}

% Introduce the topic of causal inference with Graphical Neural Networks (GNNs).

\section{Background}

% Provide background information on causal inference, neural networks, and Graphical Neural Networks (GNNs).

\section{Methodology}

% Describe the methodology of using GNNs for causal inference. Include information on how GNNs are trained, how causal inference is performed, etc.

\section{Experiments}

% Present the experiments conducted to evaluate the performance of GNNs for causal inference. Include dataset descriptions, experimental setup, and results.

\subsection{Dataset Description}

% Describe the datasets used in the experiments.

\subsection{Experimental Setup}

% Explain the experimental setup, including model architecture, training procedure, hyperparameters, etc.

\subsection{Results}

% Present and analyze the results of the experiments. Include metrics, comparisons with baseline methods, and any insights gained.

\section{Discussion}

% Discuss the implications of the results, limitations of the approach, potential future research directions, etc.

\section{Conclusion}

% Summarize the key findings of the study and conclude the paper.

\section*{Acknowledgments}

% Acknowledge any individuals or organizations that contributed to the project.

% Bibliography
\bibliographystyle{plain}
\bibliography{references} % Replace 'references' with the name of your .bib file

\end{document}